An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-09-29 poly-embed-journal/poly-embed-journal.tex
COmments from scribbles

\newcommand{\risk}[1]{\underline{#1}}
\newcommand{\inprod}[2]{\langle #1, #2 \rangle}%\mathrm{int}(#1)}
\newcommand{\relint}{\mathrm{relint}}
%\newcommand{\inter}[1]{\mathring{#1}}%\mathrm{int}(#1)}
\newcommand{\inter}{\mathrm{inter}}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\red}{\mathrm{red}}
\newcommand{\trimred}{\mathrm{trim}_\red}
\newcommand{\trimcover}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\epi}{\mathrm{epi}}
\newcommand{\indopp}{\bar{\mathbbm{1}}}
\newcommand{\ones}{\mathbbm{1}}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

@@

\section{General characteristics of polyhedra}
\subsection{Definitions and preliminaries}
\begin{definition}[Closed halfspace]
	For any $w\in\reals^d, b \in \reals$, let $H_{(w,b)}^+ := \{ x \in \reals^d \mid \inprod{x}{w} \geq b\}$ be the closed halfspace defined by $(w,b)$.
\end{definition}
\begin{definition}[Hyperplane]
	For any $(w,b)\in\reals^d \times \reals$, let $H_{(w,b)} := \{ x \in \reals^d \mid \inprod{x}{w} = b\}$ be the hyperplane generated by $(w,b)$, and the boundary of $H^+_{(w,b)}$.
\end{definition}

@@

\begin{corollary}
  A face $F_{(w,b)}$ of the polyhedron $P$ such that $F_{(w,b)} = P \cap H_{(w,b)}$ is nonempty if and only if $H_{(w,b)}$ is a supporting hyperplane of $P$.
\end{corollary}

@@

\subsubsection{Notation}
Within this appendix, we use some self-contained notation.
We will later consider losses over a finite set of outcomes $\Y$; to make notation consistent, we use $\reals^\Y_+$ to denote $\reals^{|\Y|}_+$ throughout, and let $d := |\Y|+1$.
For a fixed concave function $g : \reals^\Y_+ \to \reals_+$, define $G = \{(x,c) \mid c \leq g(x)\} \subseteq \reals^\Y_+ \times \reals$ to be the hypograph of $g$.

If $g$ takes the form of Definition~\ref{def:g-finite} below, define $H^+_v := H^+_{([v;-1], 0)}$ for a given $v$, where $[v; -1]\in\reals^\Y_+\times\reals$ define a halfspace for some $v \in \V$.
Similarly, we denote $H_y^+ := H_{([e_y; 0], 0)}^+$ for any $y \in \Y$; the latter will help us restrict our hypograph to the nonnegative orthant. 
Similarly, we let $H_v := H_{([v;-1], 0)}$ for $v \in \reals^\Y_+$ and define $H_y := H_{([e_y;0], 0)}$.

Finally, we define the face generated by $v \in \V$ with $F_v := H_v \cap G$ for any $v\in\V$.
Finally, we let $\H_{\V} = \{H_v^+ \mid v\in\V\}$ denote the set of halfspaces generated by $\V$, $\H_\Y = \{H_y^+ \mid y\in\Y\}$, $\H = \H_{\V} \cup \H_\Y$.

\subsection{Constructing a unique minimum set of facets from a finitely generated polyhedron}\label{appsubsec:phase1}
Suppose we have a finite set $\V \subset \reals_+^\Y \times \reals$.
Throughout, we will work with a function $g$ generated by $\V$ of the following form.

\begin{definition}\label{def:g-finite}
  Define the function $g : \reals_+^\Y \to \reals_+$ by
  \begin{align*}
    g(x) = \min_{v\in\V} \inprod{x}{v} - \delta(x \mid \reals_+^\Y)~,~
  \end{align*}
\end{definition}

@@

Now, we can define $G$ as the intersection of halfspaces generated by $\V$ on the nonnegative orthant.
\begin{claim}\label{claim:G-intersection-H}
  $G = \cap \H$.
\end{claim}
\begin{proof}
  $(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^\Y$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \V$ and $x_y \geq 0$ for all $y$.
  In turn, this statement holds if and only if $(x,c) \in H^+_v$ for all $v \in \V$ and in $H^+_y$ for all $y \in \Y$ respectively, so $(x,c) \in \cap \H$.
\end{proof}

We proceed with some observations about facets and dimension of the hypograph of $g$.

\begin{claim}\label{claim:G-full-dimensional}
  $G$ is $d$-dimensional.  %\raf{$g$ is nonnegative on $\reals^\Y_+$, so $G$ contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is full-dim.}
\end{claim}
\begin{proof}
  Since $g$ is nonnegative on $\reals_+^\Y$, $G$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.
\end{proof}

@@

\begin{corollary}\label{cor:anything-gen-G-subset-Hstar}
  Let $\V \subseteq \reals^\Y_+$ be a finite set such that $\H = \H_\Y \cup \H_\V$ satisfies $G = \cap \H$.  Then $\V^* \subseteq \V$.
\end{corollary}
\begin{proof}
  $\H_\Y \cup \H_{\V^*} = \H^* \subseteq \H = \H_\Y \cup \H_{\V}$, ergo $\V^* \subseteq \V$.
\end{proof}

We now show that we can equivalently construct $g$ through the unique finite set $\V^*$ instead of a generic set of vectors $\V$.
\begin{claim}\label{claim:g-gen-by-Vstar}
  $g(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals_+^\Y)$
\end{claim}
\begin{proof}
  $G = \cap \H = \cap \H^* = \cap (\H_\Y \cup \H_{\V^*}) = \{(x,c) \in \reals_+^\Y \times \reals \mid \inprod{v^*}{x} \geq c$ for all $v^* \in \V^* \}$ where the first equality follows as $\H^* \subseteq \H$.
  Since $G$ is the hypograph of $g$, this means $g$ can be written as 
  $g(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals^\Y_+)$. 
\end{proof}

@@

Suppose $L$ is a minimizable loss function.
For $x\in\reals^\Y_+$, consider the $1$-homogeneous extension of Bayes risk  $\risk{L}_+(x) := \inf_{r\in\R} \inprod{x}{L(r)}$ polyhedral.
Throughout, suppose $g(x) = \risk{L}_+(x)$.
As before, take $G$ to be the hypograph of $g$.
We define $\V = L(\R) \subseteq \reals^\Y_+$ and $\H_{\V} = \{H_v^+ \mid v\in\V\}$.  Unlike above, these could be infinite sets.  
Now let $\H = \H_\Y \cup \H_{\V}$; again, this may be infinitely generated.
\begin{claim}
	$G = \cap\H$.
\end{claim}
\begin{proof}
	Observe that $x \in \reals^\Y_+ \iff (x, c) \in \cap \H_\Y$.
	Let $x \in \reals^\Y_+$.
	\begin{align*}
	(x,c) \in G
	&\iff g(x) \geq c & \text{ since $G$ is the hypograph of $g$}\\
	&\iff \risk L_+(x) \geq c & \text{ as $g=\risk L_+$}\\
	&\iff \inprod{v}{x} \geq c \,\, \forall v \in \V & \text{By def of $\risk L_+$ as the infimum over $v \in \V$ of}\\
	& & \text{the inner product with $x$ and minimizable.} \\
	&\iff (x,c) \in H^+_{v} \,\, \forall v \in L(\R) & \text{By definition of each halfspace}\\
	&\iff (x,c) \in \cap \H_\V & \text{Since true for all $v \in \V$} \\ 
	\end{align*}
	Combining the two equalities (e.g., $\cap \H = (\cap \H_\Y) \cap (\cap \H_\V)$), we have $G = \cap \H$.

@@

%\raf{Need to use minimizable}
Take $\V^*$ to be the unique finite set such that $\H^* = \H_\Y \cup \H_{\V^*}$, where $\H^*$ is the unique minimum halfspace representation for $G$ as in \S~\ref{appsubsec:phase1}.

@@

\subsection{Projecting from $G$ to $g$}\label{subsec:project-pi}
We now consider projections from $\reals^\Y_+ \times \reals$ onto the positive orthant $\reals^\Y_+$.

>>>>

\newcommand{\risk}[1]{\underline{#1}}
\newcommand{\inprod}[2]{\langle #1, #2 \rangle}%\mathrm{int}(#1)}
\newcommand{\relint}{\mathrm{relint}}
%\newcommand{\inter}[1]{\mathring{#1}}%\mathrm{int}(#1)}
\newcommand{\inter}{\mathrm{inter}}
%\newcommand{\expectedv}[3]{\overline{#1}(#2,#3)}
\newcommand{\expectedv}[3]{\E_{Y\sim{#3}} {#1}(#2,Y)}
\newcommand{\toto}{\rightrightarrows}
\newcommand{\strip}{\mathrm{strip}}
\newcommand{\trim}{\mathrm{trim}}
\newcommand{\red}{\mathrm{red}}
\newcommand{\trimred}{\mathrm{trim}_\red}
\newcommand{\trimcover}{\mathrm{trim}}
\newcommand{\fplc}{finite-piecewise-linear and convex\xspace} %xspace for use in text
\newcommand{\conv}{\mathrm{conv}}
\newcommand{\epi}{\mathrm{epi}}
\newcommand{\hyp}{\mathrm{hypo}}
\newcommand{\indopp}{\bar{\mathbbm{1}}}
\newcommand{\ones}{\mathbbm{1}}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

@@

\section{General characteristics of polyhedra}
\jessiet{replace G with $\hyp g$ and $g := g_\V$
	break C.1.1 into its own subsection - make sure everything is completely self-contained with general functions}
\subsection{Definitions and preliminaries}
\begin{definition}[Closed halfspace]
	For any $(w,b) \in \reals^d \times \reals$, let $H_{(w,b)}^+ := \{ x \in \reals^d \mid \inprod{x}{w} \geq b\}$ be the closed halfspace defined by $(w,b)$.
\end{definition}
\begin{definition}[Hyperplane]
	For any $(w,b)\in\reals^d \times \reals$, let $H_{(w,b)} := \{ x \in \reals^d \mid \inprod{x}{w} = b\}$ be the hyperplane generated by $(w,b)$.
\end{definition} 
Observe that the hyperplane $H_{(w,b)}$ is the boundary of $H^+_{(w,b)}$.

@@

\begin{claim}
  A face $F_{(w,b)}$ of the polyhedron $P$ such that $F_{(w,b)} = P \cap H_{(w,b)}$ is nonempty if and only if $H_{(w,b)}$ is a supporting hyperplane of $P$.
\end{claim}

@@

\subsection{Notation}
Within this appendix, we use some self-contained notation.
We will later consider losses over a finite set of outcomes $\Y$; to make notation consistent, we use $\reals^\Y_+$ throughout, and let $d := |\Y|+1$.

Fix a set $\V \subseteq \reals^\Y_+$, and consider the concave function $g_\V : x \mapsto \inf_{v \in \V}\inprod{v}{x} - \delta(x \mid \reals^\Y_+)$.
We denote the hypograph of $g_\V$ by $\hyp(g_\V) = \{(x,c) \mid c \leq g(x)\} \subseteq \reals^\Y_+ \times \reals$.

Given any $v$ in a given set $\V$, define $H^+_v := H^+_{([v;-1], 0)}$, where $[v; -1]\in\reals^\Y_+\times\reals$ define a halfspace.
Similarly, we denote $H_y^+ := H_{([e_y; 0], 0)}^+$ for any $y \in \Y$; the latter will help us restrict our $\hyp(g_\V)$ to the nonnegative orthant. 
Similarly, we let $H_v := H_{([v;-1], 0)}$ for $v \in \reals^\Y_+$ and define $H_y := H_{([e_y;0], 0)}$.

%Finally, we define the face generated by $v \in \V$ with $F_v := H_v \cap G$ for any $v\in\V$.
Finally, we let $\H_{\V} = \{H_v^+ \mid v\in\V\}$ denote the set of halfspaces generated by $\V$, $\H_\Y = \{H_y^+ \mid y\in\Y\}$, $\H = \H_{\V} \cup \H_\Y$.

\subsection{Constructing a unique minimum set of facets from a finitely generated polyhedron}\label{appsubsec:phase1}
Suppose $\V$ is a finite set $\V \subset \reals_+^\Y$.
Throughout, we will work with a function $g := g_\V$ generated by $\V$ of the following form.

\begin{definition}\label{def:g-finite}
  Define the function $g_\V : \reals_+^\Y \to \reals_+$ by
  \begin{align*}
    g_\V(x) = \min_{v\in\V} \inprod{x}{v} - \delta(x \mid \reals_+^\Y)~,~
  \end{align*}
\end{definition}

@@

Throughout this section let $G:= \hyp(g_V)$ for a fixed finite set $\V$.
Now, we can define $G$ as the intersection of halfspaces generated by $\V$ on the nonnegative orthant.
\begin{claim}\label{claim:G-intersection-H}
  $G = \cap \H$.
\end{claim}
\begin{proof}
  $(x,c) \in G \iff g(x) - c \geq 0 \iff \min_{v \in \V}\inprod{v}{x} - c \geq 0$ and $x \in \reals_+^\Y$, which is true if and only if $\inprod{v}{x} - c \geq 0 \,\forall v \in \V$ and $x_y \geq 0$ for all $y$.
  In turn, this statement holds if and only if $(x,c) \in H^+_v$ for all $v \in \V$ and in $H^+_y$ for all $y \in \Y$ respectively, so $(x,c) \in \cap \H$.
\end{proof}

We proceed with some observations about facets and dimension of $G$.

\begin{claim}\label{claim:G-full-dimensional}
  $G$ is $d$-dimensional.  %\raf{$g$ is nonnegative on $\reals^\Y_+$, so $G$ contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is full-dim.}
\end{claim}
\begin{proof}
  Since $g_\V$ is nonnegative on $\reals_+^\Y$, $G$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.
\end{proof}

@@

\begin{corollary}\label{cor:anything-gen-G-subset-Hstar}
  Let $\V' \subseteq \reals^\Y_+$ be a finite set such that $\H = \H_\Y \cup \H_{\V'}$ satisfies $G = \cap \H$.  Then $\V^* \subseteq \V'$.
\end{corollary}
\begin{proof}
  $\H_\Y \cup \H_{\V^*} = \H^* \subseteq \H = \H_\Y \cup \H_{\V}$, ergo $\V^* \subseteq \V'$. \jessiet{Add refs}
\end{proof}

We now show that we can equivalently construct $g$ through the unique finite set $\V^*$ instead of the given set of vectors $\V$.
\begin{claim}\label{claim:g-gen-by-Vstar}
  $g_\V(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals_+^\Y) = g_{\V^*}(x)$
\end{claim}
\begin{proof}
  $G = \cap \H = \cap \H^* = \cap (\H_\Y \cup \H_{\V^*}) = \{(x,c) \in \reals_+^\Y \times \reals \mid \inprod{v^*}{x} \geq c$ for all $v^* \in \V^* \}$ where the first equality follows as $\H^* \subseteq \H$.
  Since $G$ is the hypograph of $g$, this means $g$ can be written as 
  $g(x) = \min_{v \in \V^*}\inprod{v}{x} - \delta(x \mid \reals^\Y_+)$. 
\end{proof}

@@

Now suppose $L$ is a minimizable loss function.
For $x\in\reals^\Y_+$, consider the $1$-homogeneous extension of Bayes risk  $\risk{L}_+(x) := \inf_{r\in\R} \inprod{x}{L(r)}$, which we assume is polyhedral throughout.
Consider the function $g(x) = \risk{L}_+(x)$.
%As before, take $G$ to be the hypograph of $g$.
We now define $\V = L(\R) \subseteq \reals^\Y_+$ and $\H_{\V} = \{H_v^+ \mid v\in\V\}$.  
Unlike above, these could be infinite sets.  
Now let $\H = \H_\Y \cup \H_{\V}$; again, this may be infinitely generated.
\begin{claim}
	$\hyp(g) = \cap\H$.
\end{claim}
\begin{proof}
	Observe that $x \in \reals^\Y_+ \iff (x, c) \in \cap \H_\Y$.
	Let $x \in \reals^\Y_+$.
	\begin{align*}
	(x,c) \in \hyp(G)
	&\iff g(x) \geq c & \text{ Definition of hypograph}\\
	&\iff \risk L_+(x) \geq c & \text{ as $g=\risk L_+$}\\
	&\iff \inprod{v}{x} \geq c \,\, \forall v \in \V & \text{By def of $\risk L_+$ as the infimum over $v \in \V$ of}\\
	& & \text{the inner product with $x$ and minimizable.} \\
	&\iff (x,c) \in H^+_{v} \,\, \forall v \in L(\R) & \text{By definition of each halfspace}\\
	&\iff (x,c) \in \cap \H_\V & \text{Since true for all $v \in \V$} \\ 
	\end{align*}
	Combining the two equalities (e.g., $\cap \H = (\cap \H_\Y) \cap (\cap \H_\V)$), we have $\hyp(g) = \cap \H$.

@@

%\raf{Need to use minimizable}
We can now use the results from \S~\ref{appsubsec:phase1} to conclude that $\hyp(g)$ has a unique minimum halfspace representation $\cap \H^*$, where $\H^* = \H_\Y \cup \H_{\V^*}$ for some unique finite set $\V^*$.
Moreover, we have $g = g_{\V^*}$, as $\H^*$ is the unique minimum halfspace representation for $\hyp(g)$.

@@

\subsection{Projecting from $\reals^d_+$ to $\reals^\Y_+$}\label{subsec:project-pi}
Suppose that we are given a set $\V$ and finite minimum representative set $\V^*$ such that $g_\V = g_{\V^*}$.\jessie{Too much assumption}

We now consider projections from $\reals^\Y_+ \times \reals$ onto the positive orthant $\reals^\Y_+$.
